Supplementary Figures for \\ Prestige bias in cultural evolutionary dynamics

### 2024-06-02 ms/si.tex
refactor si figures from ms

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title 
\title{Prestige bias in cultural evolutionary dynamics}

@@

\begin{abstract}
If the traits of more successful individuals are more likely to be adopted, the resulting cultural transmission is described as success-biased. In contrast, if the traits of ‘prestigious’ individuals–those who have already been copied many times–are more likely to be adopted, this is described as prestige-biased cultural transmission. In this case, prestige can be a convenient proxy for success. However, it is unclear how success and prestige biases interact to determine the outcome of cultural evolutionary dynamics. Here, we aim to clarify this using mathematical analysis and stochastic simulations. We find analytic approximations to the stochastic role-model choice process that facilitate the mathematical analysis and reduce the computational complexity of simulations. Approximations are given to the fixation probability and the fixation time of an invading cultural trait in different environments. Our results show that success bias effectively plays the role of natural selection, whereas prestige bias effectively plays the role of genetic drift. Prestige bias, which may be strong in highly social communities, also accelerates the evolutionary dynamics, as expected in a rich-get-richer process. These results signify a step forward in understanding how different cultural transmission biases interact.
\end{abstract}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

Cultural transmission of attitudes, preferences, beliefs, norms, and behaviors may combine vertical transmission, in which parents transmit to their offspring; oblique transmission, in which adults (teachers, leaders, and even strangers) transmit to unrelated offspring; and horizontal transmission, in which individuals from the same age cohort transmit to one another \citep{transmissionVectorsBook}.
It has been demonstrated that non-vertical cultural transmission can maintain maladaptive traits, which can be beneficial in changing environments \citep{changeEnv,Lehmann2009}.

Transmission biases may cause a cultural trait to have a higher rate of transmission than its frequency in the population.
\textit{Success bias} occurs when individuals prefer to copy from role models that demonstrate success in some activity, such as fishing, growing yams, using medicinal plants \citep{fijian_social_bias}, or hunting \citep{Mesoudi2008}, and it can increase the probability of learning a trait that is present in those successful individuals \citep{Borofsky2022}. 
Indeed, in a tournament between learning strategies  \citep{strategiesPaper}, most winning strategies included a mixture of success-biased social learning and individual learning, implying that success-biased learning is a good strategy, but that by itself it is not enough to best other strategies, even when success is measured accurately.
Jimenez and Mesoudi~\citep{prestigeSocialLearn} also note that a way to acquire adaptive social information is by preferentially copying competent individuals within a valuable domain (which they also call success bias). However, they claim that competence within a domain is often difficult or impossible to directly asses, and therefore people tend to use indirect cues of success.
Henrich and Broesch \citep{fijian_social_bias} have also suggested that direct assessment of success may be ``noisy, unreliable or unavailable'' and therefore copiers should also take into account indirect measures of perceived success (e.g., ``great fishermen may be chosen as role-models for growing yams''.)

Boyd and Richerson~\citep[Ch.~5]{evolutionBook} suggested that the assessment of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
Direct bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can try the observed paddle grips to determine which grip is better.
Frequency-dependent bias occurs when the probability of copying a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{anti-conformity} \citep{conformism}. 
The effects of conformity and anti-conformity on cultural evolution have been studied with both models \citep{Denton2020,Denton2021,Denton2022} and experiments \citep{negativeFrequency}.
Indirect bias occurs when a copier uses some observed phenotype to evaluate the attractiveness of a potential role-model. 
For example, an observer may copy the paddle grip of the ping-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
However, this may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, Boyd and Richerson~\citep[Ch. 8]{evolutionBook} suggest that maladaptive traits may spread widely in a population if indirect bias is strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient~\citep{cooperation}.

Henrich and Gil-White~\citep{prestige_evolution} noted that ``the most skilled/knowledgeable role-models will, on average, end up with the biggest and most lavish clienteles, so the size and lavishness of a given model's clientele size (the \emph{prestige}) provides a convenient and reliable proxy for that person's information quality''.
Thus, they predicted that skilled individuals have higher status, that people preferentially copy high-status individuals, and therefore that prestigious individuals may be influential even beyond their domain of expertise. 
They defined prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature \citep{prestige_evolution} .
Similarly, the New Oxford American Dictionary defines prestige as the ``widespread respect and admiration felt for someone or something on the basis of a \emph{perception} of their achievements or quality.''
Chudek et al.~\citep{prestige_cultural_learning} have also defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'' and demonstrated its occurrence in 3-4 year old children.
Nakata et al.~\citep{Nakata2024} define such a bias as a ``prestige bias based on second-order cues,'' in which ``Social learners can use the extent of attention and the amount copied from others as cues for prestige.''
Henrich and Broesch~\citep{fijian_social_bias} have further suggested that prestige bias can, over generations, lead to cultural adaptations and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits.

The distinction between success and prestige bias is important, as prestige is a context-dependent bias, rather than a content-dependent bias: it does not depend on the phenotype itself but rather on the number of copiers that have already copied each role-model, which may be easier and more accurate to estimate than success. 
Prestige bias is also frequency independent (see \Cref{cor:GBD} below), and thus it differs from conformity \citep{Denton2020,Denton2021,Denton2022}, which depends on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

Prestige bias may be more common in humans than success bias \citep{complexityPaper}.
In contemporary human society, social media make it especially easy to estimate the social and cultural influence individuals have over others, which can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices have been invented to capitalize on this effect \citep{facebook_marketing}.
However, despite many mentions of prestige in the cultural evolution literature, there are few models of prestige bias.

Here, we develop a stochastic model of cultural transmission with both indirect success bias and prestige bias to examine their relationship in contribution to the cultural evolution of populations.
We find analytic approximations for this model. We also find approximations for the probability and time to fixation of a `successful' phenotype (i.e., that is subject to success bias) in both a constant and a periodically changing environment.
Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and prestige bias play the role of natural selection and genetic drift, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Models}
We begin with a continuous-trait model with indirect success bias, previously suggested by Boyd and Richerson~\citep{evolutionBook}. Note that the indirect success bias is due to an indirect evaluation, in which a certain phenotype is used to evaluate the success of potential role-models.
We extend this model to include prestige bias, which introduces a within-generation model-choice process.
To facilitate mathematical analysis, we also develop a simpler version of the model with a dichotomous trait.

We implement our stochastic models and approximations, perform statistical analyses, and produce figures using Python \citep{python} with NumPy \citep{numpy} and Matplotlib \citep{mathplotlib}. 
Source code is available at \href{https://github.com/yoavram-lab/PrestigeBias}{https://github.com/yoavram-lab/PrestigeBias}.

\subsection*{Continuous trait}
We follow the Boyd and Richerson model~\citep{evolutionBook}, assuming only oblique transmission of a single trait. 
This focus on oblique transmission (copying from non-parental adults) means that we can neglect fitness differences between trait values since there is no correlation between the traits of parents and offspring.
We consider a population of $N$ individuals, described by a single trait that takes continuous values.
At each generation, $N$ naive individuals, or copiers, each choose a single role-model from the entire previous generation. Each copier then copies its trait value from the chosen role-model.
Note that our transmission models are slightly different from those modeled before, e.g. \citep{cumul_culture,Denton2021,evolutionBook}, in which the population is infinite and each copier samples $n$ role-models and then copies its trait from one or more of the sampled role-models.

Similar to a Wright-Fisher model, generations are non-overlapping, and the entire population is replaced in each generation.
The population at time $t$ can be described by $\vec{A}(t)=\big(A_{1}(t), \ldots, A_{N}(t)\big)$ where $A_{i}(t)$ is the trait value of individual $i$ at time $t$, and  the initial population is drawn from a standard normal distribution, $\vec{A}(0) \sim N(0,1)$ .
Cultural transmission is modeled by a function $F$ such that 
\begin{equation}\label{eq:transmission}
A_{i}(t+1) = F_i(\vec{A}(t)) \;.
\end{equation}

\paragraph{Success bias.}
Boyd and Richerson~\citep[Ch.~8, p.~247-249]{evolutionBook} describe a  transmission algorithm by defining $F$, a weighted average of the traits of all role-models, as
\begin{equation}\label{eq:boydF}
F_i(\vec{A}) = \sum_{j=1}^N G_{i,j}\cdot A_{i,j} \;, 
\end{equation}
where $G_{i,j}$ is the success bias of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:boydG}
G_{i,j} = \frac{\beta(A_{i,j})}{\sum_{k=1}^{N} \beta(A_{i,k})} \;,
\end{equation}
$A_{i,j}$ is the absolute trait value that copier $i$ estimates for role-model $j$ with some error $e_i \sim N(0,\eta^2)$,
\begin{equation}\label{eq:relativeTrait}
A_{i,j} = A_j + e_i,
\end{equation}
and $\beta(\cdot)$ is the bias function that quantifies the success bias of a role-model \citep[eq.~5.11]{evolutionBook},
\begin{equation}\label{eq:success_bias}
\beta(A_{i,j}) = b \cdot \exp{\Big(-\frac{(A_{i,j} - \hat{A})^2}{2J}\Big)} \;,
\end{equation} 
with $\hat{A}$ as the arbitrary optimal trait value, and $J$ and $b$ as parameters that control the bias strength; unless otherwise mentioned, we set $b=J=1$.
Therefore, $G_{i,j}$ is a relative success score that copier $i$ assigns to role-model $j$.

Boyd and Richerson~\citep{evolutionBook} note that the deterministic blended transmission algorithm they use has alternatives. We develop a similar stochastic model with transmission from a single random role-model where instead of \cref{eq:boydF} we define the transmission function $F$ as a random variable with its distribution given by 
\begin{equation}\label{eq:ourF}
\Pr\big(F_i(\vec{A}) = A_{j}\big) = G_{i,j} \;;
\end{equation}
{here $G_{i,j}$ is} the probability {that copier $i$ chooses} to copy the trait of role-model $j$.

\paragraph{Prestige bias.}
We introduce a new element to the model by assuming that in each generation copiers choose their role-models one by one so that the choice of one copier can affect the choice of other copiers.
We formulate this assumption in the following.
Denote by $K_{i,j}$  the number of copiers that choose role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers had already chosen a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the \emph{influence} of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The influence $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by success--the estimated biased trait value $\beta(A_{i,j})$--and prestige--the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;,
\end{equation}
where $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.
Here, the success-bias weight $\alpha_{i,j}$ determines the relative weighting of success and prestige bias. It is a characteristic of the interaction of role-model~$j$ with copier $i$ that determines the relative significance of  success vs. prestige bias in the role-model's overall influence in the eyes of the copier.
Different individuals may evaluate the importance of success and prestige differently. Additionally, we assume each role-model displays its prestige and success individually. For example, individuals with more followers but lacking skill may emphasize the number of their followers rather than their skill (i.e., have lower $\alpha_{i,j}$ value).
Finally, the trait of role-model $j$ estimated by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeTrait}.

\subsection*{Dichotomous trait}
We introduce a simplified version of the above model where the trait has only two phenotypes: an optimal phenotype $\hat{A}$ and a sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability that that phenotype is transmitted and thus prestige is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, for simplicity and for easier mathematical analysis, we assume $\alpha$ is homogeneous and constant ($\alpha_{i,j}=\alpha$), which entails exchangeability between role-models.
Therefore, the probability that the $i$-th copier {copies} phenotype $A$ is
\begin{equation}\label{eq:binary-model}
%G_{i,A} = \frac{(N-X)\alpha'\beta(A) + K_{i-1,A}}{i-1 + (N-X)\alpha'\beta(A) + \alpha'X} \;,
G_{i,A} = \frac{\alpha \cdot (N-X) \beta(A) + (1-\alpha) \cdot K_{i-1,A}}{\alpha \cdot (N-X) \beta(A) + \alpha \cdot X + (1-\alpha)\cdot(i-1)} \;,
\end{equation}
where $X$ is the number of role-models with trait $\hat{A}$ and $K_{i-1,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model, and assuming that $\beta(\hat{A})=1$ (thus the term $\alpha X$ in the denominator).
Complementing this, the probability of the $i$-th copier to copy phenotype $\hat{A}$ is $G_{i, \hat{A}} = 1-G_{i,A}$.
\Cref{fig:timeseries} shows some examples of model dynamics for various values of $\alpha$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
Our models are defined by two nested stochastic processes. Change over multiple generations is described by the {dynamics of the} phenotype distribution at each generation, $\{\vec{A}(t)\}_t$, see \cref{eq:transmission}. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$, see \cref{eq:process}.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$, where $K_{N,j}$ is the number of copiers that chose role-model $j$ after all copiers chose a role model.
However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.
Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations for the distribution of $\vec{K}_{N}$, summarized here and explained in detail below.
In both approximations we assume that the success-bias weight is either completely homogeneous, $\alpha_{i,j}=\alpha$, or that $\alpha_{i,j}=\alpha_j$ is a characteristic of role-model $j$ that does not vary between copiers.
Note that these approximations apply for both the and the continuous trait (\cref{eq:prestige}) and the dichotomous trait (\cref{eq:binary-model}) models.

\paragraph{Generalized binomial distribution approximation.}  
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore, 
(i) the expected number of copiers of role-model~$j$ equals its influence in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$); and
(ii) the expected number of copiers of each role-model equals its relative biased trait value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the bias weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated trait value.

\paragraph{Dirichlet-Multinomial distribution approximation.}
The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$). 
Hence, the number of copiers of each role-model $\vec{K_{N}}$ at the end of the role-model choice process follows the Dirichelet-Multinomial distribution.

After finding these approximations for the role-model choice process, we focus on the dichotomous-trait model, in which mathematical analysis is simpler, and studied the fixation probability and time in both a constant and a changing environment.

\subsection*{Generalized binomial distribution (GBD) approximation}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is {denoted}  by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials (the latter can be estimated from data, but its value is insignificant for our approximation). Note that $\theta=0$ gives the standard binomial distribution.
\\

\begin{result}[Generalized binomial distribution approximation]\label{res:GBD}
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the generalized binomial distribution, 
$$
K_{i,j} \sim \textit{GenBin}(i,\alpha_j\cdot\beta(A_j+e),\theta)
$$
if $e_i=e$ for all copiers $i=1, \ldots, N$; the success-bias weight only depends on the role-model and not the copier, i.e., $\alpha_{i,j}=\alpha_j$  for all $i=1, \ldots, N$; and $\theta$ is the correlation between successive role-model choices.
\end{result}

\begin{proof}
{Let} $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ {be} the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ {out} of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1) \;,
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) in \citep{GBD}, which completes the proof.
\end{proof}

This result gives the following corollary on the expected number of followers of a given role-model $j$ by the end of the role-model choice process, $K_{N,j}$.
\\

\begin{corollary}\label{cor:GBD} The expected number of copiers of role-model $j$ after all copiers have chosen a role-model is
$\E[K_{N,j}] = N \cdot G_{1,j}$, where $G_{1,j}$ is the probability of the first copier to copy role-model $j$.
In addition, $\E[K_{N, j}] = \alpha_j \cdot \beta(A_j+e) / \overline{\alpha \cdot \beta(A+e)}$, where the averaging in the denominator is over the role-models index, $j$.
\end{corollary}

\begin{proof}
The expected value of the GBD is
$\E[K_{N,j}] = N \cdot Q_j(1,1)$, see Drezner and Farnum~\citep[eq.~(2.3)]{GBD}.
Here, $Q_j(1,1)$ is the initial probability to choose role-model $j$, before any role-model choices are made, such that $Q_j(1,1) = G_{1,j}$ by definition.
The rest of the proof is in Appendix~\ref{sec:GBD}.
\end{proof}

Note that $G_{1,j} = \alpha_{i,j} \beta(A_{i,j}) / \sum_{i=1}^{N}{\alpha_{i,j} \beta(A_{i,j})}$ (see \cref{eq:prestige} with $K_{0,j}=0$). In the limit of $\alpha_{i,j} \to 0$, that is with only prestige bias, we get $G_{1,j} = 1/N$, and from \Cref{cor:GBD}, the expected number of copiers of role-model $j$ is 1. Therefore, prestige bias is frequency independent, in contrast to conformity bias.

The special case where the bias weight is uniform for all role-models ($\alpha = \alpha_j$ for $j=1,\ldots, N$) is interesting, because we can evaluate the expected number of copiers using a linear equation
\begin{equation}\label{eq:GBDsamelalpha}
\E[K_{N,j}]= N\cdot\frac{\alpha\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N}\alpha\cdot\beta(A_m+e)} =\beta(A_j+e) \bigg/ \overline{\beta(A+e)} \;,
\end{equation}
where the only variable is $A_j+e$, because $\overline{\beta(A+e)}$ is the mean of the distribution of the trait values, modified by some constant parameters of $\beta$.
We can then write $L = 1/\overline{\beta(A+e)}$ and
\begin{equation}\label{eq:linearEq}
\E[K_{N,j}] = L\cdot \beta(A_j+e) \;.
\end{equation}

\paragraph{Numerical validation.}
To validate that the GBD approximation for the number of copiers of a role-model is correct (\cref{eq:GBDsamelalpha}), 
we ran $1,000$ simulations of the full model, and compared the results with \Cref{cor:GBD}.
We compare the distribution of the number of copiers by plotting the histograms of both our simulation results and the expected values based on \Cref{cor:GBD}.

Although basic, \Cref{fig:GBD} shows a good fit of the GBD approximation.
We perform more extensive validations on the Dirichlet-Multinomial approximation (see below), because that is what we will use in our analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dirichlet-Multinomial distribution (DMD) approximation}

\paragraph{P\'{o}lya urn model.}
This stochastic process consists of $N$ draws from an urn with an initial {number} of colored balls of $M$ colors. When a ball is drawn, it is then placed back in the urn together with an additional new ball of the same color.
Let $\vec{U_i} = \{u_{i,1},u_{i,2},...,u_{i,M}\}$  where $u_{i,j}$ is the number of balls of the $j$-th color in the urn after $i$ draws.
Let $S_{i,j}=1$ when drawing a $j$-colored ball on the $i$-th draw, and $0$ otherwise. The probability that $S_{i,j}=1$ given $\vec{U_{i-1}}$ is
\begin{equation}\label{eq:polya}
\begin{split}
P(S_{i,j} = 1 \mid \vec{U_{i-1}}) = 
\frac{u_{i-1,j}}{\sum\limits_{m=1}^{M} u_{i-1,m}} = 
\frac{o_j + w_{i-1,j}}{\sum\limits_{m=1}^{M} o_m + w_{i-1,m}} = 
\frac{o_j + w_{i-1,j}}{i-1 + \sum\limits_{m=1}^{M} o_m} \;,
\end{split}
\end{equation}
where $o_j$ is the initial number of balls of color $j$ in the urn, and $w_{i,j}$ is the cumulative number of new balls that were added to the urn after $i$ draws of color $j$.
\\

\begin{result}[P{\'{o}}lya urn model]\label{result:polya}
The role-model choice process, $\big\{\vec{K}_i\big\}_{i=1}^N$, is equivalent to a \emph{P\'{o}lya urn model} if both the trait estimation error and the success-bias weight are constant in the population, $e_i=e$ for all $i=1, \ldots, N$ and $\alpha_{i,j}=\alpha$  for all $i,j=1, \ldots, N$.
\end{result}

\begin{proof} 
{Write} $\alpha'=\frac{\alpha}{1-\alpha}$ as the success-bias weight ratio, and $A'_j=A_j+e$. From \cref{eq:prestige} and because $\sum_{j=1}^{N}{K_{i,j}}=i$, we have
\begin{equation}\label{eq:copier_choose}
G_{i,j} = 
\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{\sum\limits_{m=1}^{N}{\left[ \alpha'\beta(A'_m) + K_{i-1,m}\right]}}
 =\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{i-1 + \sum\limits_{m=1}^{N}\alpha'\beta(A'_m)} \;.
\end{equation}
Substituting $M=N$, $o_j = \alpha'\beta(A'_j)$, and $w_{i,j} = K_{i,j}$ in \cref{eq:polya} gives \cref{eq:copier_choose}, thus completing the proof.
\end{proof}

Frigyik et al.~\citep[section 2]{dirichlet} prove that the proportion of different colored balls in a \emph{P\'{o}lya urn model} converges to the Dirichlet distribution as the number of draws  (the population size, $N$, in our model) approaches infinity, based on the \emph{Martingale Convergence Theorem} \citep{martingaleBook}.
We therefore have an approximation for the relative influence each role-model has when evaluated by copiers. Thus, choosing the role-models for all copiers is equivalent to drawing from a {multinomial} distribution where the parameters are the modified weights from a Dirichlet distribution and we have the following corollary.
\\

\begin{corollary}[Dirichlet-Multinomial distribution approximation]\label{cor:dirichlet}
The number of copiers of each role-model approximates a Dirichlet-Multinomial distribution, $\vec{K_N} \sim \textit{DirMul}(N,\vec{G_1})$, under the conditions of \Cref{result:polya}.
\end{corollary}

\paragraph{Numerical validation.}
We next validated the DMD approximation of our model
and tested its sensitivity to the assumptions ($e_i=e$ and $\alpha_i=\alpha$ for $i=1,\ldots,N$) by comparing results of stochastic simulations of our model (\cref{eq:binary-model}) with the DMD approximation (\Cref{cor:dirichlet}).
We used a relatively small population size, $N=100$, thus validating that the approximation is in good agreement even for small $N$, despite the assumption of large $N$ in the proof by Frigyik et al.~\citep[section 2]{dirichlet}.
First, we computed an observed distribution of the number of copiers from the average empirical distribution of multiple simulations.
We then compared this observed distribution with the expected theoretical DMD (\Cref{fig:DM_validation}A).
The difference in distributions was not perceived when plotting both distributions on the same figure, so we used the difference instead.
The maximum difference is 0.5 role-models, which indicates a very good fit.
In addition, we tested the likelihood of the observed data to be drawn from the DMD, against a shuffle of the parameters vector of the DMD itself (\Cref{fig:DM_validation}B). We see that the negative log likelihood of the observed data is much higher than any other shuffled version of the parameters vector, strongly supporting our approximation.

Next, we examined the fixation probability and fixation time of a `successful' phenotype $\hat{A}$ (i.e., that is favored by success bias) when invading a population of phenotype $A$ and compared results from the full model and the DMD approximation.
Thus, we assume the population has a single individual with phenotype $\hat{A}$ and $N-1$ individuals with phenotype $A$. 
We find that the number of simulations needed to sufficiently approximate our model with the DMD approximation is roughly $1,000$ (\Cref{fig:num_sims}).
We examined the robustness of the DMD approximation to relaxing the following assumptions.
First, we relaxed our assumption of no estimation error $e$.
Estimation error in the original model was drawn from a normal distribution, and added to the trait value before evaluation of the success `bias ($A_{i,j} = A_j + e_i$).
When estimation error is applied, we sample $e_i$ for each copier $i$ from a normal distribution with expected value zero and variance $\eta^2$.
Even when the estimation error variance is $0.1$, both the fixation probability and fixation time DMD approximations are accurate~(\Cref{fig:hetro_error}). 
We also relaxed our assumption of a uniform bias weight $\alpha$ (i.e., $\alpha_i=\alpha$). We allowed $\alpha$ to vary in the population, drawing $\alpha_j$ for each role-model $j$ from a normal distribution such that $\alpha_j \sim N(0.5, \epsilon)$ where $\epsilon$ is between $10^{-7}$ and $10^{-1}$. 
We found again that results of the DMD approximation are similar to those from stochastic simulations of the full model~(\Cref{fig:hetro_alpha}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fixation probability and time}
After finding that the DMD is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and conditional fixation time (conditioned on the population reaching fixation) of a `successful' phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the success-bias weight, $\alpha$, which determines the relative effect of success and prestige bias, given by \cref{eq:prestige}.

For simplicity, we use the dichotomous-trait model, which also assumes a constant success-bias weight $\alpha_{i,j}=\alpha$, and we do not include trait estimation error in this analysis, i.e., $e_i=0$.
As shown above, transmission in our model is approximately Dirichlet-Multinomial distributed (\Cref{cor:dirichlet} and \cref{eq:copier_choose}).
We focus on two scenarios: the first scenario is of a `constant environment' in which the same phenotype, $\hat{A}$, is always favored by success bias; the second scenario is of a `changing environment' in which the phenotype favored by success bias cycles between the invading phenotype $\hat{A}$ and the resident phenotype $A$ (i.e., $\hat{A}$ starts as the rare phenotype).

\paragraph{Drift and diffusion terms in a constant environment.}
We start by finding the expectation and variance of the change in frequency from one generation to the next, which are the drift and diffusion terms of the diffusion equation.
Let $x$ and $x'$ be the frequency of phenotype $\hat{A}$ in a population with $N$ individuals in the current and next generation, respectively.
We set $\beta$ to be the success bias of phenotype $A$ relative to phenotype $\hat{A}$, such that $\beta = \beta(A)/\beta(\hat{A}) < 1$.
Then (see Appendix~\ref{sec:drift_diff_const} for derivation),
\begin{equation}\label{eq:drift_diff}
\begin{aligned}
E[x'-x] &= x(1-x)(1-\beta) + o(1-\beta)  \;, 
\\
V(x'-x) &= x(1-x)\left(\frac{1}{\alpha N + (1-\alpha)}\right) + o\left(\frac{1}{\alpha N + (1-\alpha)}\right) \;.
\end{aligned}\end{equation}

This analysis gives an interesting result relating the parameters $\alpha$ and $\beta$ to parameters of the classical Wright-Fisher model from population genetics:
the selection coefficient $s$, a measure of the effect of natural selection on the change in frequency of genotypes, and the effective population size, $N_e$, a measure of the effect of random genetic drift on the change in frequency of genotypes. 
In a diffusion-equation approximation of the classical Wright-Fisher model, the expectation and variance of the change in frequency are $E[x'-x]=sx(1-x)+o(s)$ and $V[x'-x]=x(1-x)/N_e$ \citep[eq.~7]{kimura}{, respectively}. 
Therefore, we have the following result.\\

\begin{result}[Effective selection coefficient and population size]\label{res:selection_coef}
The effective selection coefficient $s$ and effective population size $N_e$ can be written in terms of the success coefficient $\beta$ (\cref{eq:success_bias}), the success-bias weight $\alpha$ (\cref{eq:prestige}), and the population size $N$ as 
\begin{equation}\label{eq:effective}
s=1-\beta=\frac{\beta(\hat{A})-\beta(A)}{\beta(\hat{A})}, \quad N_e=\alpha N + (1-\alpha) \;.
\end{equation}
\end{result}

Note that when $N>>1$, $N_e \approx \alpha N$, resulting in a very convenient expression.\\

Using our effective selection coefficient, $s=1-\beta$, and effective population size, $N_e$, with the population-genetics fixation probability approximation given by Kimura~\citep[eq.~8]{kimura}, we obtain the following result.\\

\begin{result}[Fixation probability]
The fixation probability of an invading phenotype favored by success bias is approximately 
\begin{equation}\label{eq:kimura_p}
\pi(x) = \frac{1-e^{-2(1-\beta)N_e x}}{1-e^{-2(1-\beta)N_e}},
\end{equation}
where $x$ is the initial frequency of the invading phenotype.
\end{result}

Similarly, we can use $1-\beta$ and $N_e$ in the population-genetics fixation time approximation given by \citep[eq.~17]{kimura_average}.\\

\begin{result}[Fixation time] The expected fixation time (conditioned on fixation) from an initial frequency $x$ is approximately
\begin{equation} \label{eq:kimura_t}
T(x) = J_1(x) + \frac{1-\pi(x)}{\pi(x)}\cdot J_2(x),
\end{equation}
where $N_e=\alpha N + (1-\alpha)$, $S=N_e(1-\beta)$, and
\begin{equation}
\begin{aligned}
J_1(x) &= \frac{-1}{(1-\beta)(e^{-2S}-1)}\int_x^1 \frac{1-e^{2S\xi}-e^{-2S(1-\xi)}+e^{-2S}}{\xi(1-\xi)}d\xi \;, \\
J_2(x) &= \frac{-1}{(1-\beta)(e^{-2S}-1)}\int_0^x \frac{(1-e^{2S\xi})(e^{-2S\xi}-1)}{\xi(1-\xi)}d\xi \;.
\end{aligned}
\end{equation}
\end{result}
Note that these integrals cannot be solved in closed form, and are estimated numerically.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title 
\title{Supplementary Figures for \\ Prestige bias in cultural evolutionary dynamics}

@@

\beginsupplement % https://support.authorea.com/en-us/article/how-to-create-an-appendix-section-or-supplementary-information-1g25i5a/

\begin{figure}[h]
    \includegraphics[width=1\linewidth]{timeseries.pdf}
  \caption{
  \textbf{Example dynamics of the dichotomous model with varying bias weight.}
  The dichotomous model is defined in eq.~11. Shown are full dynamics without DM or GBD approximations.
  Here, population size, $N=1,000$; ideal phenotype value, $\hat{A}=1$; success bias value, $\beta(A)=0.99$; initial phenotype frequencies are 50\%-50\%.
  }	
  \label{fig:timeseries}
\end{figure}
